Optimizing and Incrementalizing Collection Queries by AST Transformations

### 2018-01-29 aosd13/pepm13/caseStudyDiscussion.tex
Automated search and replace of Sec.~\ref with \cref
Fix/Cleanup setting of TeX-master
It didn't work for some files in aosd13.
Fix position of footnote mark

%One such context is:
%\begin{lstlisting}
%def asExp[T](t: Exp[T]) = t
%\end{lstlisting}
%Note that \code{asExp} is an identity function, but restricted to expression trees; this triggers the Scala compiler to use an implicit resolution if needed. Standard rules for implicit resolution apply, which select the implicit conversion with the most specific type. Hence, while \code{pure((a, b))} would produce \code{Const((a, b))} with type \code{Exp[(Exp[A], Exp[B])]}, 
%\code{asExp((a, b))} produces \code{LiftTuple2(a, b)} with type \code{Exp[(A, B)]}, that is a representation of \code{(a, b)}, as desired.
%
%
%\pg{Actually, the problem shows up with any two nested constructors of this kind.}
%
%EDSL programs need to resort to calling \code{asExp} extremely rarely, because operations of the EDSL already provide the needed context; but as soon as we try supporting nesting tuples such as \code{((a, b), c}, we face again type inference bugs, for which this time we found no reasonable workaround which does not affect surface syntax. The conversion above accepts a pair of expression trees; we can write a more general conversion which accepts a pair of values \emph{implicitly convertible} to expression trees, so that \code{asExp((a, b), c)} would have type \code{Exp[((A, B), C)]}, but the compiler cannot make use of this conversion\footnote{We reported \url{https://issues.scala-lang.org/browse/SI-5651}, probably a dup of \url{https://issues.scala-lang.org/browse/SI-3346}.}.
%The problem generalizes to any two nested generic constructors.
%
%We observe that these limitations of type inference make its behavior hard to predict and hinder DSL embedding. We discuss this point further in Sec.~\ref{sec:ScalaLessons}.

@@

\smartParagraph{Overloading}
The code in Fig.~\ref{fig:reifiedQuery} uses the lifted \code{BookData} constructor. Two definitions of \code{BookData} are available, with signatures \code{BookData(String, String, Int)} and \code{BookData(Exp[String], Exp[String], Exp[Int])}, and it seems like the Scala compiler should be able to choose which one to apply using overload resolution. This however is not supported simply because the two functions are defined in different scopes\footnote{\url{https://issues.scala-lang.org/browse/SI-2551}}, hence importing \code{BookData(Exp[String], Exp[String], Exp[Int])} shadows locally the original definition.
%\pg{Resume}
% function of type \code{}
% are reported as bugs; some of them were already acknowledged, others were found to be already known.
%During our work we reported many of the bugs we found; we provide links to the ones which are still not addressed and which support 
%\pg{Say something about bugs found.}

@@

Because of this type inference limitations, we failed to provide support for reifying code like \code{((a, b), c)}\footnote{One could of course write a specific implicit conversions for \emph{this} case; however, \code{(a, (b, c))} requires already a different implicit conversion, and there are infinite ways to nest pairs, let alone tuples of bounded arity.}.
%Hence, we did not find a way to support nested tuples like

@@

% vim: set ft=tex:

>>>>

%One such context is:
%\begin{lstlisting}
%def asExp[T](t: Exp[T]) = t
%\end{lstlisting}
%Note that \code{asExp} is an identity function, but restricted to expression trees; this triggers the Scala compiler to use an implicit resolution if needed. Standard rules for implicit resolution apply, which select the implicit conversion with the most specific type. Hence, while \code{pure((a, b))} would produce \code{Const((a, b))} with type \code{Exp[(Exp[A], Exp[B])]},
%\code{asExp((a, b))} produces \code{LiftTuple2(a, b)} with type \code{Exp[(A, B)]}, that is a representation of \code{(a, b)}, as desired.
%
%
%\pg{Actually, the problem shows up with any two nested constructors of this kind.}
%
%EDSL programs need to resort to calling \code{asExp} extremely rarely, because operations of the EDSL already provide the needed context; but as soon as we try supporting nesting tuples such as \code{((a, b), c}, we face again type inference bugs, for which this time we found no reasonable workaround which does not affect surface syntax. The conversion above accepts a pair of expression trees; we can write a more general conversion which accepts a pair of values \emph{implicitly convertible} to expression trees, so that \code{asExp((a, b), c)} would have type \code{Exp[((A, B), C)]}, but the compiler cannot make use of this conversion\footnote{We reported \url{https://issues.scala-lang.org/browse/SI-5651}, probably a dup of \url{https://issues.scala-lang.org/browse/SI-3346}.}.
%The problem generalizes to any two nested generic constructors.
%
%We observe that these limitations of type inference make its behavior hard to predict and hinder DSL embedding. We discuss this point further in \cref{sec:ScalaLessons}.

@@

\smartParagraph{Overloading}
The code in Fig.~\ref{fig:reifiedQuery} uses the lifted \code{BookData} constructor. Two definitions of \code{BookData} are available, with signatures \code{BookData(String, String, Int)} and \code{BookData(Exp[String],}\linebreak\code{Exp[String], Exp[Int])}, and it seems like the Scala compiler should be able to choose which one to apply using overload resolution. This however is not supported simply because the two functions are defined in different scopes,\footnote{\url{https://issues.scala-lang.org/browse/SI-2551}} hence importing \code{BookData(Exp[String], Exp[String], Exp[Int])} shadows locally the original definition.
%\pg{Resume}
% function of type \code{}
% are reported as bugs; some of them were already acknowledged, others were found to be already known.
%During our work we reported many of the bugs we found; we provide links to the ones which are still not addressed and which support
%\pg{Say something about bugs found.}

@@

Because of these type inference limitations, we failed to provide support for reifying code like \code{((a, b), c)}\footnote{One could of course write a specific implicit conversions for \emph{this} case; however, \code{(a, (b, c))} requires already a different implicit conversion, and there are infinite ways to nest pairs, let alone tuples of bounded arity.}.
%Hence, we did not find a way to support nested tuples like

@@
